A brief study of Approximation Algorithms for 2D Packing Problem

### 2021-12-10 paper.tex
Fix grammar errors in the first 3 sections

This problem is very important in many fields. In the transportation industry, this algorithm can be used to pack an item into containers. There is also an interesting application of this problem in main memory allocation in multi-processor programming. Basically, this problem can be adapted to most resource allocation problems with two dimension resource constraints.\par

@@

We already show that 2D Packing Problem to be NP-Complete, and P vs NP remains an open problem, so we can not find an efficient algorithm to deterministically solve it. We must try another approach. To solve an inherently difficult problem, various approaches have been well-studied. For decision problems, we can use randomized algorithms to give the correct result with high probability. For optimization problems, we can use another approach: Approximation algorithm\cite{garey1976approximation} which is guaranteed to produce a sub-optimal but reasonable result. \par
In this paper, we are going to investigate various approximation algorithms for the 2D Packing Algorithm. In section 2, we will briefly study the definition and performance metric for Approximation algorithms. In section 3, we will introduce two early level-oriented heuristic approximation algorithm\cite{coffman1980performance} for 2D Packing Problem. In section 4,  we will study a more advanced approach\cite{steinberg1997strip}.

@@

The analysis for Approximation algorithms often requires a thorough mathematical proof for the worst-case performance. We can find some surprising properties for hard problems through Approximation algorithms. Although all NP-Complete problems are equivalently difficult, their complexity varies from the perspective of Approximation algorithms. For example, the Knapsack problem can produce solutions that are arbitrarily close to the optimal result. However, finding a solution that is constant-bounded to the optimal result for the Maximum clique problem is proved to be equivalent to prove whether P = NP. Therefore, Algorithm provided an extra classification for NP-hard problems.\par

@@

Then for minimization problems, Approximation ratio $\alpha$ puts a multiplicative maximum bound on the result $A(L) \leq \alpha*OPT(L)$. For maximization, approximation ratio puts a multiplicative minimum bound on the result $A(L) \geq \frac{OPT(L)}{\alpha}$. We can also call this ratio Absolute Approximation Ratios to be precise.

@@

Different from Approximation ratio that only uses a multiplicative bound factor, Asymptotic Approximation uses an additional additive bound factor. This gives us a more fine-grained metric to measure the performance of Approximation algorithms. For minimization problem, an Asymptotic bound $\alpha$ means that:

@@

NFDH proceed as follows: After we sort all the rectangles with decreasing height. We simply pack rectangles left-justified from left to right on the bottom of the bin (the first level), so the first rectangle is the highest rectangle. When there is no more sufficient width in the current level, we proceed to the next level (a horizontal line on top of the first rectangle in the current level). We just iterate from level to level until we pack all the rectangles.

@@

We will also provide an intuition into the complicated proof: Measuring the approximation ratio for 2D Packing is equivalent to measure how much space we are wasting during packing. For the NFDH algorithm, the wasted space is the empty space between levels. By drawing a horizontal line of height $H_{i+1}$ (which is $\leq H_i$) at level $i$, we are able to partition the wasted space into 2 areas $S_i$ (the empty space above the line) and $T_i$ (the empty space below the line). $S$ has width at most $1$ and height at most $H_i - H_{i+1}$, $T$ has width $1-W_i$ which is smaller that $x_{i+1}$ accoring to the inequality $W_{i} + x_{i+1} \geq 1$, $T$ has height $H_{i+1}$. Then we know that $S_i \leq 1*(H_i - H_{i+1}) = H_i - H_{i+1}$ and $T_i \leq (1 - W_i)*H_{i+1} \leq x_{i+1}*H_{i+1} \leq A_{i+1}$. Then by summing them up, we get $$S = \sum_{i=1}^{t} S_{i} \leq \sum_{i=1}^{t} (H_i - H_{i+1}) = H_1 - H_t \leq H_1 = h_{max}$$ $$T = \sum_{i=1}^{t}T_i \leq \sum_{i=1}^{t}A_{i+1} \leq A$$
Then we are only wasting $A + h_{max}$ spaces, therefore the result: $$NFDH(L) \leq A+S+T \leq 2*A + h_{max} \leq 2*OPT(L) + h_{max}$$
It is also worth mentioning that $S$ (the space between the top of rectangles and the level lines) is the total space wasted by using a level-based approach rather than putting every rectangle at the lowest possible position. Because $S \leq h_{max}$, we know that the level-based approach will only introduce a maximum waste of an additive constant and will not affect the asymptotic approximation ratio.

@@

\paragraph{Recursive construction for $M$} Now we turn to the construction for the core algorithm $M$ for the correctness of the algorithm above.

>>>>

This problem is very important in many fields. In the transportation industry, this algorithm can be used to pack items into containers. There is also an interesting application of this problem in main memory allocation for multi-processor programming. Basically, this problem can be adapted to most resource allocation problems with two dimension resource constraints.\par

@@

We already show that 2D Packing Problem is NP-Complete, and P vs NP remains an open problem, so we can not find an efficient algorithm to deterministically solve it. We must try another approach. To solve an inherently difficult problem, various approaches have been well-studied. For decision problems, we can use randomized algorithms to give the correct result with high probability. For optimization problems, we can use another approach: Approximation algorithm\cite{garey1976approximation} which is guaranteed to produce a sub-optimal but reasonable result. \par
In this paper, we are going to investigate various approximation algorithms for the 2D Packing Algorithm. In section 2, we will briefly study the definition and performance metric for Approximation algorithms. In section 3, we will introduce two early level-oriented heuristic approximation algorithms\cite{coffman1980performance} for 2D Packing Problem. In section 4,  we will study a more advanced approach\cite{steinberg1997strip}.

@@

The analysis for Approximation algorithms often requires a thorough mathematical proof for the worst-case performance. We can find some surprising properties for hard problems through Approximation algorithms. Although all NP-Complete problems are equivalently difficult, their complexity varies from the perspective of Approximation algorithms. For example, there exist an algorithm for the Knapsack problem that can produce solutions that are arbitrarily close to the optimal result. However, finding a solution that is constant-bounded to the optimal result for the Maximum clique problem is proved to be equivalent to prove whether P = NP. Therefore, Approximation algorithms offer an extra, more fine-grained classification for NP-hard problems.\par

@@

Then for minimization problems, Approximation ratio $\alpha$ puts a multiplicative maximum bound on the result $A(L) \leq \alpha*OPT(L)$. For maximization problems, approximation ratio puts a multiplicative minimum bound on the result $A(L) \geq \frac{OPT(L)}{\alpha}$. We can also call this ratio Absolute Approximation Ratios to be precise.

@@

Different from Absolute Approximation ratio that only uses a multiplicative bound factor, Asymptotic Approximation uses an additional additive bound factor. This gives us a more fine-grained metric to measure the performance of Approximation algorithms. For minimization problems, an Asymptotic bound $\alpha$ means that:

@@

NFDH proceeds as follows: After we sort all the rectangles with decreasing height. We simply pack rectangles left-justified from left to right on the bottom of the bin (the first level), so the first rectangle is the highest rectangle. When there is no more sufficient width in the current level, we proceed to the next level (a horizontal line on top of the first rectangle in the current level). We just iterate from level to level until we pack all the rectangles.

@@

We will also provide an intuition into the complicated proof: Measuring the approximation ratio for 2D Packing is equivalent to measure how much space we are wasting during packing. For the NFDH algorithm, the wasted space is the empty space between levels. For each level $i$, we draw a horizontal line of height $H_{i+1}$ (which is $\leq H_i$), then we are able to partition the wasted space into 2 areas $S_i$ (the empty space above the line) and $T_i$ (the empty space below the line). $S_i$ has width at most $1$ and height at most $H_i - H_{i+1}$. Because we are packing rectangles with decreasing height, every rectangles in level $i$ is higher than $H_{i+1}$, therefore every $T_i$ has width $1-W_i$ which is smaller than $x_{i+1}$ according to the inequality $W_{i} + x_{i+1} \geq 1$, $T_i$ has height $H_{i+1}$. Then we know that $S_i \leq 1*(H_i - H_{i+1}) = H_i - H_{i+1}$ and $T_i \leq (1 - W_i)*H_{i+1} \leq x_{i+1}*H_{i+1} \leq A_{i+1}$. Then by summing them up, we get $$S = \sum_{i=1}^{t} S_{i} \leq \sum_{i=1}^{t} (H_i - H_{i+1}) = H_1 - H_t \leq H_1 = h_{max}$$ $$T = \sum_{i=1}^{t}T_i \leq \sum_{i=1}^{t}A_{i+1} \leq A$$
Then we are only wasting $A + h_{max}$ spaces, therefore the result: $$NFDH(L) = A+S+T \leq 2*A + h_{max} \leq 2*OPT(L) + h_{max}$$
It is also worth mentioning that $S$ (the wasted space between the top of rectangles and the level lines) is the total space wasted by using a level-based approach rather than putting every rectangle at the lowest possible position. Because $S \leq h_{max}$, we know that the level-based approach will only introduce a maximum waste of an additive constant and will not affect the asymptotic approximation ratio.

@@

\paragraph{Recursive construction for $M$} Now we turn to the construction for the core algorithm $M$ for the correctness of the algorithm above. Steinberg designed several mutually disjoint reduction procedures that are applied when specific geometric properties are observed in input instances. Specifically, the input is categorized to fall into one of four classes $C_0, C_1, C_2, C_3$, where corresponding procedures $P_0,P_1, P_2,P_3$ would either solve $L$ or reduce to problems of smaller sizes.
\par The four classes are designed as below.\\
(C0) $w_M\leq \frac{w}{2}, h_M\leq\frac{h}{2},$ and for some index $i\in[1,n]$ we have $A-\frac{1}{4}wh\leq a_i$; \\
(C1) $w_M\geq \frac{w}{2}$; \\
(C2) $w_M\leq \frac{w}{2}, h_M\leq\frac{h}{2}$, and there exists two different indices $i$ and $k$ such that $w_i\geq \frac{w}{4}, w_k\geq\frac{w}{4}, h_i\geq\frac{w}{4}, h_k\geq\frac{h}{4}, 2(A-a_i-a_k)\leq (w-\max(w_i, w_k))h$; \\
(C3) $w_M\leq\frac{w}{2}, h_M\leq\frac{h}{2}, n>1$, and for some index $m\in[1,n)$ we have $A-\frac{1}{4}wh\leq\sum_{i=1}^{m}s_i\leq\frac{3}{8}wh$ and $w_{m+1}\leq\frac{w}{4}$. 
\par By switching the positions of $w_M$ with $h_M$ and $w$ with $h$, the above conditions can be further expanded to $C_{-1}, C_{-2}, C_{-3}$ which allows for the rotation of any rectangles. It is shown that if input instance $(Q,L)$ satisfies inequalities \ref{cbase}, then it must satisfy one of the conditions $C_\mu$ with $\mu\in[-3,3]$ \cite{steinberg1997strip}.
\par For input instances satisfying C0 apart from inequalities \ref{cbase}, P0 first places rectangle $R_i$ so that $_*[R_i]=_*Q$, where $_*Q$ denotes the left bottom corner of bin $Q$. If $n=1$, then $(Q,L)$ is solved; otherwise, P0 forms a new problem $(Q^\prime, L^\prime)$, where new bin $Q^\prime$ is defined by
	_*[Q^\prime]=[R_i]_*, [Q^\prime]^*=Q*
where $[R_i]_*$ denote the right bottom corner for $R_i$ and $Q^*$ denotes the upper right corner for $Q$. $L^\prime$ is obtained by removing $R_i$ from $L$. This input 
